Covariate Analysis of ROSMAP Reprocessed counts Regressing out Diagnosis as well for TWAS

### 2019-11-08 code/ROSMAP_RegressDiagnosis_4_TWAS_hg38.Rmd
Mayo hg38

syn3382527.7
syn4300313.1
syn8449369.2
syn7116000.1
#_#OLDCOUNTS syn8691134.1
syn21075911

#SynID Translator
syn21075913

syn3191087.3
#_#OLDMetrics syn8698240.2 Version 2
syn21075912 
#^^^ Version 1

@@

GENE.PARAM = syn_temp$get('syn8449369', version = 2)$path %>%
  data.table::fread(data.table = FALSE)
ALL_USED_IDs = c(ALL_USED_IDs, 'syn8449369')
GENE.LEN = dplyr::select(GENE.PARAM, ensembl_gene_id, gene.length) %>% 
  unique() 
rownames(GENE.LEN) = GENE.LEN$ensembl_gene_id
GENE.GC = dplyr::select(GENE.PARAM, ensembl_gene_id, percentage_gc_content) %>% 
  unique() 
rownames(GENE.GC) = GENE.GC$ensembl_gene_id 
```

@@

### Filter genes
* Remove genes that have less than 1 cpm counts in at least 50% of samples per Diagnosis
* Remove genes with missing gene length and percentage GC content
```{r filter.genes}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
  }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$ensembl_gene_id[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$gene.length)]) %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
# Check gene biotype
## Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "dec2016.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
## Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "gene_biotype"),
                       filters = "ensembl_gene_id", 
                       values = PROCESSED_COUNTS$filteredExprMatrix$genes,
                       mart = mart)
summary(factor(Ensemble2HGNC$gene_biotype)) %>%
  rownameToFirstColumn('Biotype') %>%
  dplyr::rename(fraction = DF) %>%
  dplyr::mutate(fraction = fraction/dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]) %>%
  dplyr::filter(fraction >= 0.01) %>%
  kable
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

@@

Syns_Used <- c("syn3382527", "syn4300313", "syn8449369", "syn7116000",
  "syn21075911", "syn3191087", "syn21075913", "syn21075912"
)
# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  tissueTypeAbrv	= 'DLPFC', 
  study = 'ROSMAP', 
  organism = 'HomoSapiens',
  consortium	= 'ROSMAP',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh37'
)
# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'ROSMAP_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_SVA_Diagnosis_Corrected.tsv', name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

>>>>

syn3382527.7
syn4300313.1
----Old Lengths syn8449369.2
syn21156161
syn7116000.1
----OLDCOUNTS syn8691134.1
syn21075911
----SynID Translator
syn21075913
syn3191087.3
----OLDMetrics syn8698240.2 Version 2
syn21075912

@@

GENE.PARAM = syn_temp$get('syn21156161', version = 2)$path %>%
  data.table::fread(data.table = FALSE)
ALL_USED_IDs = c(ALL_USED_IDs, 'syn21156161')
GENE.LEN = dplyr::select(GENE.PARAM, ensembl_gene_id, gene.length) %>% 
  unique() 
rownames(GENE.LEN) = GENE.LEN$ensembl_gene_id
GENE.GC = dplyr::select(GENE.PARAM, ensembl_gene_id, percentage_gc_content) %>% 
  unique() 
rownames(GENE.GC) = GENE.GC$ensembl_gene_id 
```

@@

### Filter genes
* Remove genes that have less than 1 cpm counts in at least 50% of samples per Diagnosis
* Remove genes with missing gene length and percentage GC content
```{r filter.genes}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                     MIN_GENE_CPM=1, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
    processed.counts$filteredExprMatrix$genes
  }, COUNT)
genesToAnalyze = unlist(genesToAnalyze) %>% 
  unique() %>%
  intersect(GENE.GC$ensembl_gene_id[!is.na(GENE.GC$percentage_gc_content)]) %>%
  intersect(GENE.LEN$ensembl_gene_id[!is.na(GENE.LEN$gene.length)]) %>%
  setdiff(c("N_unmapped", "N_multimapping", "N_noFeature", "N_ambiguous"))
PROCESSED_COUNTS = getGeneFilteredGeneExprMatrix(COUNT[genesToAnalyze, ], 
                                                 MIN_GENE_CPM=0, 
                                                 MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0)
# Check gene biotype
## Define biomart object
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", host = "sep2019.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
## Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol", "gene_biotype"),
                       filters = "ensembl_gene_id", 
                       values = PROCESSED_COUNTS$filteredExprMatrix$genes,
                       mart = mart)
summary(factor(Ensemble2HGNC$gene_biotype)) %>%
  rownameToFirstColumn('Biotype') %>%
  dplyr::rename(fraction = DF) %>%
  dplyr::mutate(fraction = fraction/dim(PROCESSED_COUNTS$filteredExprMatrix$genes)[1]) %>%
  dplyr::filter(fraction >= 0.01) %>%
  kable
```
Processing `r dim(PROCESSED_COUNTS$filteredExprMatrix)[1]` genes in `r dim(PROCESSED_COUNTS$filteredExprMatrix)[2]` samples

@@

Syns_Used <- c("syn3382527", "syn4300313", "syn21156161", "syn7116000",
  "syn21075911", "syn3191087", "syn21075913", "syn21075912"
)
# Set annotations
all.annotations = list(
  dataType = 'mRNA',
  dataSubType = 'geneExp',
  summaryLevel = 'gene',
  assay	 = 'RNAseq',
  tissueTypeAbrv	= 'DLPFC', 
  study = 'ROSMAP', 
  organism = 'HomoSapiens',
  consortium	= 'ROSMAP',
  normalizationStatus	= TRUE,
  normalizationType	= 'CQN',
  rnaquantification = 'RSEM',
  genomeAssemblyID = 'GRCh37'
)
# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'ROSMAP_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ROSMAP_SVA_Diagnosis_Corrected.tsv', name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), used = Syns_Used, activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
